---
title: "Term Project"
author: "The Big Apple group"
date: "November 4, 2017"
output: html_document
---
```{r}
ibm.data= read.csv("C:/Manthans Work/GitHub/Staff Attrition/HRAttrition.csv", header= T)
```

\n
# To view data sets and summary to get information about distribution
```{r}
#View(ibm.data)
head(ibm.data)
summary(ibm.data)
```
\n
# Exploring variable to collect information

```{r}
pie=ggplot(ibm.data , aes(x=factor(1), fill=Attrition))+
  geom_bar(width = 1)+
  coord_polar("y")
pie

```
```{r}
round(prop.table(table(ibm.data$Attrition)),digit=2)

```


```{r}
table(ibm.data$Education, ibm.data$EducationField)
#table(ibm.data$EmployeeNumber)
```

```{r}
hist(ibm.data$JobSatisfaction)
abline(v = mean(ibm.data$JobSatisfaction), col = "blue", lwd = 2)

#Average job satisfaction of employees is 2.7
```


\n
#Converting the type of variables into factor and vieweing summary of the end result. 
```{r}
ibm.data$Education<- as.factor(ibm.data$Education)
ibm.data$JobSatisfaction<- as.factor(ibm.data$JobSatisfaction)
ibm.data$NumCompaniesWorked<- as.factor(ibm.data$NumCompaniesWorked)
ibm.data$PerformanceRating<- as.factor(ibm.data$PerformanceRating)
ibm.data$RelationshipSatisfaction<- as.factor(ibm.data$RelationshipSatisfaction)
ibm.data$StockOptionLevel<- as.factor(ibm.data$StockOptionLevel)
ibm.data$WorkLifeBalance<- as.factor(ibm.data$WorkLifeBalance)


ibm.data$EnvironmentSatisfaction<- as.factor(ibm.data$EnvironmentSatisfaction)
ibm.data$JobInvolvement<- as.factor(ibm.data$JobInvolvement)
ibm.data$JobLevel<- as.factor(ibm.data$JobLevel)


```


```{r}
dplyr::glimpse(ibm.data)
```
\n
#Information related to NA values
```{r}
sum(is.na(ibm.data))
```
\n
#Plotting histogram to get information related to monthly income and years at company. 
```{r}
#hist(ibm.data$DailyRate)
#hist(ibm.data$HourlyRate)
yes_month=ibm.data[ibm.data$Attrition == "Yes" & ibm.data$Department == "Sales",]
hist(yes_month$MonthlyIncome)
#hist(ibm.data$MonthlyRate)
hist(ibm.data$YearsAtCompany)



```

\n
#Exploring all the major variables in the data sets. 
```{r}
#prop.table(table(ibm.data$Attrition,ibm.data$BusinessTravel))
#prop.table(table(ibm.data$Attrition,ibm.data$BusinessTravel), margin = 1)
round(prop.table(table(ibm.data$Attrition,ibm.data$BusinessTravel), margin = 1), digits = 2)


#table(ibm.data$Attrition,ibm.data$Department)
#prop.table(table(ibm.data$Attrition,ibm.data$Department), margin = 2)
round(prop.table(table(ibm.data$Attrition,ibm.data$Department), margin = 2), digits = 2)

#table(ibm.data$Attrition,ibm.data$EnvironmentSatisfaction)
#round(prop.table(table(ibm.data$Attrition,ibm.data$EnvironmentSatisfaction), margin = 2), digits = 2)

#table(ibm.data$Attrition,ibm.data$Gender)
round(prop.table(table(ibm.data$Attrition,ibm.data$Gender), margin = 2), digits = 2)

round(prop.table(table(ibm.data$JobInvolvement)), digits = 2)
#table(ibm.data$Attrition,ibm.data$JobInvolvement)
round(prop.table(table(ibm.data$Attrition,ibm.data$JobInvolvement), margin = 2), digits = 2)

#table(ibm.data$Attrition,ibm.data$JobLevel)
#round(prop.table(table(ibm.data$Attrition,ibm.data$JobLevel), margin = 2), digits = 2)

round(prop.table(table(ibm.data$Attrition,ibm.data$JobSatisfaction),margin = 2), digits = 2)

#table(ibm.data$Attrition,ibm.data$MaritalStatus)
round(prop.table(table(ibm.data$Attrition,ibm.data$MaritalStatus),margin = 1), digits = 2)

#table(ibm.data$Attrition,ibm.data$OverTime)
round(prop.table(table(ibm.data$Attrition,ibm.data$OverTime),margin = 2), digits = 2)

#table(ibm.data$Attrition,ibm.data$RelationshipSatisfaction)
round(prop.table(table(ibm.data$Attrition,ibm.data$RelationshipSatisfaction),margin = 2), digits = 2)

#table(ibm.data$Attrition,ibm.data$WorkLifeBalance)
round(prop.table(table(ibm.data$Attrition,ibm.data$WorkLifeBalance),margin = 2), digits = 2)


```
\n
#Plotting the histogram to provide information between monthly income and attrition. 

```{r}
head(ibm.data)
library('lattice')

histogram(~JobSatisfaction | Attrition, data = ibm.data)
histogram(~Department | Attrition, data = ibm.data)
histogram(~JobLevel | Attrition, data = ibm.data)
#histogram(~BusinessTravel | Attrition, data = ibm.data)
#histogram(~DistanceFromHome | Attrition, data = ibm.data)
#histogram(~Education | Attrition, data = ibm.data)
#histogram(~EducationField | Attrition, data = ibm.data)
histogram(~EnvironmentSatisfaction | Attrition, data = ibm.data)
#histogram(~Gender| Attrition, data = ibm.data)
##histogram(~HourlyRate | Attrition, data = ibm.data)
#histogram(~JobInvolvement | Attrition, data = ibm.data)
histogram(~JobRole | Attrition, data = ibm.data)
histogram(~MaritalStatus | Attrition, data = ibm.data)
#histogram(~NumCompaniesWorked | Attrition, data = ibm.data)
histogram(~OverTime | Attrition, data = ibm.data)
#histogram(~PercentSalaryHike | Attrition, data = ibm.data)
#histogram(~PerformanceRating | Attrition, data = ibm.data)
histogram(~RelationshipSatisfaction | Attrition, data = ibm.data)
histogram(~TotalWorkingYears | Attrition, data = ibm.data)
#histogram(~WorkLifeBalance | Attrition, data = ibm.data)
histogram(~YearsAtCompany | Attrition, data = ibm.data)
histogram(~YearsInCurrentRole | Attrition, data = ibm.data)
#histogram(~YearsSinceLastPromotion | Attrition, data = ibm.data)
histogram(~YearsWithCurrManager | Attrition, data = ibm.data)
```
```{r}
histogram(~EnvironmentSatisfaction, data = ibm.data)
mean(as.numeric(ibm.data$EnvironmentSatisfaction))
```


\n
#Plotting boxplot to get information related to age and monthly income. 
```{r}
#boxplot(ibm.data$MonthlyIncome, main = "Income")
#boxplot(ibm.data$ï..Age, main = "Age")
bwplot(~Age, ibm.data)
bwplot(~MonthlyIncome, ibm.data)
```

\n
#Table to relate worklifebalance and marital status of the employees who left the organization. 
```{r}
prop.table(table(ibm.data$Attrition, ibm.data$WorkLifeBalance, ibm.data$MaritalStatus))
```

\n
#Below code is used to calculate information related to Research and development department employee who left the organization considering distancefromhome and overtime variable to get relationship. 
```{r}
ibm_rd_left = (ibm.data[ibm.data$Department=="Research & Development" & ibm.data$Attrition == 'Yes',])
hist(ibm_rd_left$DistanceFromHome)
ibm_RD = (ibm.data[ibm.data$Department=="Research & Development" & ibm.data$Attrition == 'Yes' & ibm.data$OverTime == 'Yes', ])

head(ibm_RD)
hist(ibm_RD$DistanceFromHome)
boxplot(ibm_RD$DistanceFromHome)
```

```{r}

```


```{r}
set.seed(1234)
indicies=sample(1:2 , length(ibm.data$Age) , replace = T, prob=c(.8,.2)) 


ibm.train =ibm.data[indicies==1,]
ibm.test  =ibm.data[indicies==2,] 

```

```{r}
#View(ibm.train)
```


```{r}
ibm.model= glm(Attrition~Department+EnvironmentSatisfaction+Gender+JobLevel+JobRole+JobSatisfaction+MaritalStatus+OverTime+TotalWorkingYears+RelationshipSatisfaction+YearsAtCompany+YearsInCurrentRole+YearsSinceLastPromotion+YearsWithCurrManager, family = "binomial",data=ibm.train)
summary(ibm.model)
```

```{r}
#This package requies JDK files in Program Files to run

#install.packages("rJava")
#install.packages("glmulti")
#library(glmulti)
require('glmulti')
```














```{r}
ibm.glmulti.model <-
    glmulti(Attrition ~Department+EnvironmentSatisfaction+Gender+JobLevel+JobRole+JobSatisfaction+MaritalStatus+OverTime+TotalWorkingYears+RelationshipSatisfaction+YearsAtCompany+YearsInCurrentRole+YearsSinceLastPromotion+YearsWithCurrManager,data= ibm.train,
            level = 1,               # interaction considered
            method = "h",            # genetic approach
            crit = "aic",            # AIC as criteria
            confsetsize = 2,         # Keep 2 best models
            plotty = F, report = F,  # No plot or interim reports
            fitfunction = "glm",     # glm function
            family = binomial)       # binomial family for logistic regression
```


```{r}
## Show 2 best models (Use @ instead of $ for an S4 object)
ibm.glmulti.model@formulas
```
```{r}
summary(ibm.glmulti.model@objects[[1]])
#summary(ibm.glmulti.model@objects[[2]])
```



```{r}
ibm.pmodel=ibm.glmulti.model@objects[[1]]

my.predictions=predict(ibm.pmodel,ibm.test, type='response' )
predicted.classes=as.integer(ifelse(my.predictions>0.5, 1,0))
table(predicted.classes, ibm.test$Attrition)


```

```{r}
contrasts(ibm.test$Attrition)
```

```{r}
table(ibm.test$Attrition)
round(prop.table(table(ibm.test$Attrition)),digits = 2)
```

```{r}
#True Accuracy
8/(31+8)
```


```{r}
#Model Accuracy

(251+8)/296
```


